### 2006-03-28 doc/paper/skysubtraction.tex
concurrent update

\subsection{Extraction}

We utilize an optimal extraction procedure to transform a 2-dimensional image of
the fiber spectrograms to a set of 320 1-dimensional spectra (per spectrograph).
We use the same procedure on all processed frames, 
and will describe the details in processing the flat-field frames, the
arc lamp frames, and finally the science frames.  
The extraction is performed row-by-row, so each frame requires an 
outside loop over 2048 individual rows.
The counts in each row of 2048 columns are modeled with a linear 
combination of 320 profiles plus a low-order
polynomial background.  In addition, deviations to the trace centers and
the profile widths are calculated as linear basis modes (reflecting the first
and second spatial derivatives, respectively).  The compiled routine
we implement to fit the profiles in a single row (extract_row.c) can
be invoked with any combination of these parameters.

We decided to perform profile fitting to the 
SDSS spectral images for the following reasons:

\begin{enumerate}

\item{The fiber profiles overlap significantly with their nearest neighbors.}

\item{The profiles are relatively stable and smoothly vary over 
location on the CCDs.  They result from the convolution of the fiber 
illumination pattern + the point-spread function
of the optics in the SDSS spectrographs + the window funciton of the 
Site CCD 24 micron square pixels. }

\item{A least-squares fit with the true profile delivers an 
unbiased and minimum variance estimate of the counts (e.g. Horne et al. 1995).}

\end{enumerate}

The major obstacle to actual optimal extraction is the determination of 
the exact shape and position of the underlying profile.  
We investigated a range of symmetric, analytic profiles with a 
minimum number of free parameters.  In the end, we found the best 
results using a simple Gaussian or a slightly steeper profile 
characterized by $e^{-|x|^3/\sigma^3}$.  The profiles are defined to be 
normalized to unity, so the amplitude of the profile fit is a 
direct estimator of the counts at a single spectral position.

Extraction is first carried out on all flat-field exposures 
(typically 10 second observations of 4 quartz lamps projected on petals covering the entrance aperture of the telescope) which were observed
with a common plug plate configuration.  The flat-field extraction is 
used to determine the shape of
the profile as a function of wavelength and position on the CCD.  
The resulting spectral extractions are also used to 
determine the empirical fiber-to-fiber variations in throughput 
(as a function of wavelength) during a single setup.   
A flat-field frame is labeled rejected by the pipeline if one of the
following conditions is met:  1) More the 2\% of the CCD pixels are labeled
bad by sdssproc, 2) More than 100 rows of the image show saturation, 
3) The 80th percentile count level is less than 1000 counts, 
4) Fewer than all 8 flat-field telescope petals were closed, 
5) Fewer than all 4 flat-field lamps were on,
6) More than 10000 pixels were rejected during fiber tracing, 
7) Fiber tracing fails, or traces are seperated by less than 3 pixels anywhere
on the flat-field image.

The first very critical step is to robustly determine the fiber positions 
in CCD pixel coordinates.
This is performed by identifying peaks in a cross-sectional cut of across 
the center row (actually a median cross-section of the central 9 rows).  
The correct identification of the 320 fibers per frame can be 
uniquely determined by demanding a wider spacing between the bundles 
of 20 fibers (spaced by about 10 pixels) relative to the nominal inter-fiber 
spacing (6.1 pixels on average).  Once the starting peaks and 
identification have been determined, we trace the fiber centers 
by performing flux-weighted centroids on subsequent rows
using the previous row's center as the intial guess.  This has proven to 
work well on SDSS flat-fields, and we have yet to uncover a failure 
mode in over 10000 flat-field frames.
The flux-weighted centroids do have scatter from shot noise, 
so we fit a 7th order polynomial to each set of fiber centroids, 
and the best-fit coefficients are stored to represent the fiber positions 
on each flat-field frame.

Each flat-field frame is extracted twice, once assuming a Gaussian
profile and again assuming the steeper $|x|^3$ profile.  In both cases,
the width is allowed to vary freely.  The total number of free parameters
per row in the flat-field extractions is equal to 2 per fiber plus 10
parameters to represent a smooth polynomial background.  After the extractions
are complete, $\chi^2$ is computed by directly comparing the model fit
to the flat-field image.  The median $\chi^2$ for each case is compared
directly, and the better fit and profile (lower median $\chi^2$) are retained
and implemented for the extraction of the remaining frames.  Also retained
is a smooth polynomial fit the average profile width in each fiber bundle
as a function of CCD row number.

After all flat-field frames for a given sequence have been extracted
and stored, the arc-lamp frames in the same sequence are extracted.  
Each arc-lamp frame is matched with the nearest (in time) 
non-rejected flat-field frame.  The matched flat-field gives the 
initial trace centers to
be assumed, as well as the profile shape and width determined from the
flat-field.  To account for any flexure between the frames, the trace
centers are allowed to shift/stretch to the arc-lamp image given a 
3x3 polynomial distortion kernel over the entire extent of the image
(match_trace.pro).

The arc-lamp frames are extracted row-by-row, in the same
manner as the flat-fields, the number of free parameters assumed
for the model fit is 2 per fiber (profile area and width) and 1 parameter
to fit for a constant background per row.  The restricted freedom in the
background parameterization reflects the lower scattered light levels in
the images of the arc-lamps.

\subsection{(or arc-lamp spectra -> wavelength map)}

We determine an initial wavelength solution for every arc-lamp frame
successfully extracted in the following manner:

For the appropriate side of each spectrograph (red or blue), we 
construct a list of suitable unblended lines from a parent list including
known lines of Mercury, Cadmium, Neon and (a trace of) Argon 
that are readily observable in the standard set of SDSS arc-lamps. 
A representative spectrum is calculated as the median extracted spectrum
of the 5 closest "unflagged" spectra falling near the central columns 
of the arc-lamp image.  This median arc-spectrum is matched with 
cross-correlation to a simulated spectrum of suitable arc-line positions and 
strengths by systematically searching a small volume of 
parameter space which conservatively covers
the polynomial coefficients of the SDSS wavelength solution.
In the first iteration, we search a set of 
100 coefficient vectors allowing for quadratic changes and return 
the best match.  In the second iteration, we search in a smaller region
about the previous best match, and search a list of 125 coefficient vector
allowing for cubic deviations.  The coefficient vector with the best
match from the second iteration is returned as an initial guess along with
the cross-correlation coefficient.

A 2-step iterative procedure is loop through
a range of possible polynomial

\begin{enumerate}
\item{Locate bright fibers (aka whopping fibers)}
\item{Assess sky fiber levels}
\item{Commence iterative Optimal extraction }
\item{Shift wavelength solution to match sky lines (describe in wavelength solution)}
\item{Sky subtract}

\end{enumerate}

Note: Flux correction and Telluric correction have been moved (a long time ago).

\subsection{Sky Subtraction}

Sky subtraction is performed on a frame by frame basis using the 
"flattened" wavelength calibrated extractions.  The goal of sky subtraction
is to estimate the expected background counts contributing to each of 
the native 320x2048 pixels in each extracted frame.  A "supersky" model 
at a fiducial airmass of 1 is constructed by performing iterative fits to
the extracted spectra associated with blank sky positions (so-called
"sky" fibers).   The following steps are taken in 3 sequential iterations
in order to produce the sky-subtracted spectra:

\begin{enumerate}

\item{Identify the "sky" fibers which have no bits flagged 
in the fiber mask array.}

\item{Divide each sky spectrum by the airmass of the fiber position appropriate
for the midpoint of the observation to produce respresentative sky spectra for 
an airmass of 1.  The inverse variance is scaled accordingly.}

\item{Group and sort the sky spectra as a function of wavelength.  This step
transforms a 2-dimensional set of N_sky x 2048 pixels into a sorted 1-dimensional array.  
Each pixel maintains its associated central wavelength and inverse variance. }

\item{Perform a single cubic b-spline fit as a function of wavelength.  
The number of breakpoints (number of free parameters is 4 greater) is set 
at 2048 on the red side and 1536 on the blue side.  The breakpoint spacing 
is set automatically to maintain approximately constant S/N between breakpoint 
positions. The b-spline fit itself is iterative, with upper and lower rejection 
thresholds set to mask bad or deviant pixels. }

\item{The b-spline fit is evaluated at all 320x2048 wavelengths representing all
spectral pixels in the frame.  The airmass correction is applied for each individual 
fiber, and the rescaled super-sky is subtracted from each spectrum.}

\item{Estimate the significance of the residuals in the sky fibers used for the
fit in wavelength bins separated by breakpoints determined above.  for each bin
where inverse the 67th percentile of $\chi^2$ is greater than 1, we rescale 
all the inverse variance values down by the 67th percentile value of $\chi^$.
The rescaling is actually done with a smooth interpolation as a function of 
redshift with a constraint that the inverse variances cannot increase. }

\end{enumerate}

%The b-spline routines from the numerical fortran library: {\bf slatec}
%were ported to IDL to be easily incorporated into the pipeline.

The sequence followed in the determination of the sky subtracted spectrum
and re-scaled errors is called 3 times successively.  After the 1st call,
any sky fibers which have a median \chi^2 per pixel greater than 2 is 
masked as a 'BADSKYFIBER' and the corresponding bit is set in the fibermask
array.  If any fibers are masked, the sky subtraction is iterated a 2nd time.
The sky subtraction routine is invoked for one final iteration, but with extra 
freedom allowed in the fit.  The same freedom is allowed in wavelength, by
keeping the same number of breakpoints, but a quadratic polynomial fit is
allowed as a function of wavelength.  This triples the number of free 
parameters in the linear least-squares fit, but the values are still 
overconstrained due to the ample number of sky fibers.  The extra freedom
in the model can accomodate variation in the PSF of sky features as a function 
of position on the CCD.

The sky subtraction method which utilizes the b-spline model takes advantage 
of the inherent over-sampling in wavelength due to the large number of 
sky fibers assigned per plate, without the requirement of resampling the 
pixel value to a fixed linear wavelength scale.  The least squares fit
is carried out for native spectral pixels where the errors are very nearly 
uncorrelated and the fit is evaluated for all the native pixels (320x2048)  
at the central wavelengths of the single spectral frames.

The accuracy of the SDSS spectral sky subtraction has been tested in recent
scientific analyses using large numbers of background-limited spectra.   
The distribution of flux residuals between spectral observations 
on separate nights has been shown to be 8\% larger in width than the average expected
from the reported errors (McDonald et al. 2004, Burgess \& Burles 2005).
The spectra of faint, background-limited galaxies observed with SDSS exhibit a 
distribution of residuals about the best fit spectral models described by the sum of a
Gaussian and Laplace distribution (Bolton et al. 2004).  
The influence of the exponential tail is greatly enhanced in the vicinity of strong sky 
emission lines representative of systematic errors in the brightest lines, but the gaussian 
width closely is closely represented by the reported variance array.